---
title: "A comparison of motivation playlists"
author: "Tessa Vermeulen"
date: 'February/March 2022'
output:
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: cerulean
    self_contained: false
---

```{r setup}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidymodels)
library(plotly) # N.B. Requires the Cairo package
library(protoclust)
library(heatmaply)
library(spotifyr)
library(compmus)
library(scales)
library(gridExtra)
library(grid)

```
### Long-distance and sprinters playlists compared using the __tempo__ feature with standard deviation and mean

```{r sd}
long_dist <-
  get_playlist_audio_features(
    "thesoundsofspotify",
    "56vEaKq9VqFBlqtcXZAgn4"
  ) %>%
  slice(1:30) %>%
  add_audio_analysis()
sprinters <-
  get_playlist_audio_features(
    "thesoundsofspotify",
    "70wUKUusJoxhLNieGGRXpK"
  ) %>%
  slice(1:30) %>%
  add_audio_analysis()
jazz <-
  long_dist %>%
  mutate(genre = "Long-Distance") %>%
  bind_rows(sprinters %>% mutate(genre = "Sprinters"))


jazz %>%
  mutate(
    sections =
      map(
        sections,                                    # sections or segments
        summarise_at,
        vars(tempo, loudness, duration),             # features of interest
        list(section_mean = mean, section_sd = sd)   # aggregation functions
      )
  ) %>%
  unnest(sections) %>%
  ggplot(
    aes(
      x = tempo,
      y = tempo_section_sd,
      colour = genre,
      alpha = loudness
    )
  ) +
  geom_point(aes(size = duration / 60)) +
  geom_rug() +
  theme_minimal() +
  ylim(0, 5) +
  labs(
    x = "Mean Tempo (bpm)",
    y = "SD Tempo",
    colour = "Playlist",
    size = "Duration (min)",
    alpha = "Volume (dBFS)"
  )

```

***
Looking at the plot, we can see that in general, the tempo of the songs in the sprinters playlist are higher. 
For now it is a bit difficult to see, but I also would like to say something about the volume. It looks like the volume of the sprinters playlist is higher for most songs, but I think it is better if I showed a little table with the mean volume of both playlists. 

### Can we say something about the __keys__ of the different playlists?

#### Key histogram

```{r key_hist}
long_distance <- get_playlist_audio_features("","56vEaKq9VqFBlqtcXZAgn4")
sprinters <- get_playlist_audio_features("", "70wUKUusJoxhLNieGGRXpK")


total <-
  bind_rows(
    long_distance %>% mutate(category = "Long Distance swimmers"),
    sprinters %>% mutate(category = "Sprinters")
  )

key_histogram <- total %>%
  ggplot(aes(x=key, fill=category)) +
  geom_histogram(position = 'dodge', binwidth = 0.55) +
  theme_classic()

ggplotly(key_histogram)
```

*** 
The histogram of the keys shows what keys are used in the different playlists. As we can see, the majority of the tracks of the sprinters playlist ar in key 1 (C#/Db). This corresponds to the fact that the sprinters playlist mainly contains rock songs. For the long distance playlist, the distribution is more spread out. 

Furthermore, the sprinter playlist does not have any tracks that are in key 7 (G), and the Long-Distance playlist does not have tracks in key 11 (B)

#### Keygram of the song "Nothing else matters"

```{r keygram}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )


nothing_else_matters <-
  get_tidy_audio_analysis("0nLiqZ6A27jJri2VCalIUs") %>%
  compmus_align(sections, segments) %>%
  select(sections) %>%
  unnest(sections) %>%
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      )
  )

#keygram
nothing_else_matters %>% 
  compmus_match_pitch_template(
    key_templates,         # Change to chord_templates if descired
    method = "euclidean",  # Try different distance metrics
    norm = "manhattan"     # Try different norms
  ) %>%
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")

```

***
The song Nothing else matters was the song with the lowest energy. So lets take a closer look at this song. 
A keygram of a song compares  the keys of a song to a particular key template, to figure out the key of a song. When we look at the plot, the x-axis shows the time and duration of the song, and the y-axis shows the different possible keys. A dark color in the plot means that key matches the key in the template. If the color gets lighter, there isn't a key that matches a particular key template. 

So at the beginning of the song, and towards the end, there are no good matches. However, around time 100s we can see a very dark color for the E:min chord, so chances are that this chord matches the template at this time in the song. 

Notes: 
- ik wil graag de uitleg naast de plots ipv eronder

### Self-similarity matrix of Shivers from Ed Sheeran (timbre)

```{r self-sim}
shivers <-
  get_tidy_audio_analysis("50nfwKoDiSYg8zOCREWAm5") %>% # Change URI.
  compmus_align(bars, segments) %>%                     # Change `bars`
  select(bars) %>%                                      #   in all three
  unnest(bars) %>%                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

ed_sheeran <- 
  shivers %>%
    compmus_self_similarity(timbre, "euclidean") %>% 
    ggplot(
      aes(
        x = xstart + xduration / 2,
        width = xduration,
        y = ystart + yduration / 2,
        height = yduration,
        fill = d
      )
    ) +
    geom_tile() +
    coord_fixed() +
    scale_fill_viridis_c(guide = "none") +
    theme_classic() +
    labs(x = "", y = "")


ggplotly(ed_sheeran)

```

***
The song Shivers from Ed Sheeran is one of the songs in my personal playlist. This song falls under the Pop genre. Most pop songs contain an intro, verses, pre-chorus and chorus. 
With a self-similarity matrix, we can determine which sections of a song more or less have the same timbre, or same overall sound, and which sections are different. 
When looking closely at the plot, we can see two diagonal lines off the main diagonal. This means that a section of the song has been repeated. In pop songs, this is usually an indication of a repetition of the chorus.

* plot mooier opmaken
* plot verder analyseren

### Sprinters vs Long-distance swimmers 

#### Introduction
Let me introduce myself. My name is Tessa, and I am a professional athlete. I swim in the national team and have won several national and international medals. 

During competitions I listen to music that inspires and motivates me. So, for me and a lot of my team mates, music plays an important role in the optimal race preperation. 

In swimming there are different diciplines. More specificly there are sprinters, and long-distance swimmers. Sprinters exert maximum effort for a short period of time, often no longer than 30 seconds. A long-distance swimmer does not have to be so explosive, and the exertion of their race usually lasts more than two minutes. 

Of course not all swimmers lissten to music during their preperantion, but in my experience there are both sprinters and long-distance swimmers who have special playlists for competitions. 


#### Research question

Since sprinters and long distance differ so much from each other in terms of performances, the research question I will try to answer in this portfolio is: 

**What are the differences between the motivation playlists of sprinters and long-distance swimmers?**

First we look at the track level Spotify features of both playlists. Are there for example differences in energy, or tempo? I expect that the sprinter playlist contains more songs with higher energy and tempo. 

#### Playlists 

For this portfolio, I used several playlists of my teammates. I have subdivided these playlists in two groups, namely a playlist with tracks that sprinters listen to and one that long-distance swimmers listen to. 

**Typical tracks for sprinters:**

* lakdjal
* lajdjla

**Typical tracks for long-distance swimmers:**

* asda
* ;kjakdf


### pitch chromatogram
```{r pitch_chromatogram}
# Leas on life andy grammer
andy_grammer <-
  get_tidy_audio_analysis("4ufkuONjQMNR2fyXu1bO9w") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)

#all I do is win dj Khaled
dj_khaled <-
  get_tidy_audio_analysis("12PNcnMsjsZ3eHm62t8hiy") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)

both <- 
  bind_rows(
    andy_grammer %>% mutate(category = "Andy Grammer - Leas on life"),
    dj_khaled %>% mutate(category = "DJ Khaled - All I do is win")
  )

pitch <- 
  both %>%
    mutate(pitches = map(pitches, compmus_normalise, "euclidean")) %>%
    compmus_gather_chroma() %>% 
    ggplot(
      aes(
        x = start + duration / 2,
        width = duration,
        y = pitch_class,
        fill = value
      )
    ) +
    geom_tile() +
    labs(x = "Time (s)", y = NULL, fill = "angular") +
    theme_minimal() +
    facet_wrap(~category)




ggplotly(pitch)

```

### Changes in __energy__ and __danceability__ point out the differences between sprinters and long-distance swimmers

```{r energy}
long_distance <- get_playlist_audio_features("","56vEaKq9VqFBlqtcXZAgn4")
sprinters <- get_playlist_audio_features("", "70wUKUusJoxhLNieGGRXpK")


total <-
  bind_rows(
    long_distance %>% mutate(category = "Long Distance swimmers"),
    sprinters %>% mutate(category = "Sprinters")
  )

e <- ggplot(total, aes(category, energy)) + 
  geom_boxplot()

#ggplotly(e, tooltip = c("category"))

d <- ggplot(total, aes(category, danceability)) + 
  geom_boxplot()
#ggplotly(d, tooltip = c("category"))

grid.arrange(e, d, ncol=2)
#ggplotly(plot)

```

***
As the plots show, sprinters listen to music with higher energy, and the songs of long-distance swimmers have higher danceability. One outlier in energy is the song "Nothing else matters" from Metallica. This song is ... lower in energy than them mean of the long-distance playlist. 
A reason for this could be that people listen to this song because they want to control their nerves before the race and dont want to listen to a very energetic song. Another reason that this song is in this playlist, is that lyrics of the song is also an important factor.

#### What __genres__ are used in both playlists? 

```{r genres}

```

***
the main genre of the sprinters playlist is rock
the main genre of the long-distance playlist is pop 

### Conclusion 

Still working on the conclusion 